---
title: "Project1"
author: "Nihar Dave"
date: "2023-02-21"
output: word_document
---

```{r}
library(fpp3) 
library(tidyverse) 
library(readxl) 
library(readr) 
library(dplyr) 
library(ggplot2)
library(forecast)

library(readr) 
library(stringr) 
library(knitr) 
library(seasonal) 
library(forecast) 
```

```{r}
setwd("/Users/nihardave/Desktop/Sem2/Statistical Forecasting")
library(readxl)

Project1_data <- read_excel("Project1_data.xls")%>%
  mutate(Date = as.Date(Date), 
         Unemployment_Rate = as.numeric(Unemployment_Rate)) %>%
  as_tsibble(index = Date)
Project1_data

```

#### Creating time series object:

```{r}
data_ts <- ts(Project1_data$Unemployment_Rate, start = c(1982,12,01), frequency =12)
```

#### Plotting Time series:

```{r}
autoplot(data_ts) + xlab("Time") + ylab("Unemployment Rate") + ggtitle("Unemployment Rate of Canada aged 15 or more")
```

#### ACF plot

```{r}
acf(data_ts,lag.max = 120)
ggtsdisplay(data_ts)
```

#### Decompose time series:

```{r}
data_decomp <- decompose(data_ts)
autoplot(data_decomp)
```

#### Training and testing data set:

```{r}
train_data <- window(data_ts, end = c(2017,12))
test_data <- window(data_ts, start = c(2017,01))
```

1.  Fitting a ETS model & generating forecast for testing data:

```{r}
data_ets <- ets(train_data)
data_fc <- forecast::forecast(data_ets, h = length(test_data))
```

Plotting forecast & actual values:

```{r}
autoplot(data_fc) + xlab("Time") + ylab("Unemployment Rate") + ggtitle("Unemployment rates Forecast using ETS model") +
  autolayer(test_data, series = "Actual") + 
  guides(colour = guide_legend(title = "Forecast"))
```

2.  Fitting a Mean model & generating forecast for testing data:

```{r}
mean <-  meanf(data_ts, h=12)
mean_forecast <- forecast::forecast(mean, h=12)
```

Plotting forecast & actual values:

```{r}
autoplot(mean_forecast) + xlab("Time") + ylab("Unemployment Rate") + ggtitle("Unemployment rates Forecast by mean model") +
  autolayer(test_data, series = "Actual") +
  guides(colour = guide_legend(title = "Forecast"))
```

3.  Fitting a Naive model & generating forecast for testing data:

```{r}
naive_data <-  naive(data_ts)
naive_forecast <- forecast::forecast(naive_data)
```

Plotting forecast & actual values:

```{r}
autoplot(naive_forecast) + xlab("Time") + ylab("Unemployment Rate") + ggtitle("Unemployment rates Forecast by naive model") +
  autolayer(test_data, series = "Actual") +
  guides(colour = guide_legend(title = "Forecast"))
```

4.  Fitting a STL model & generating forecast for testing data:

```{r}
data_stl <- stlf(data_ts, s.window = "periodic")

```

Plotting forecast & actual values:

```{r}
autoplot(data_stl) + xlab("Time") + ylab("Unemployment Rate") + ggtitle("Unemployment rates Forecast by STL model") +
  autolayer(data_ts, series = "Actual") + 
  guides(colour = guide_legend(title = "Forecast"))
```

1.  Residual for ETS forecasting:

```{r}
ets_residual <- residuals(data_ets)
autoplot(data_ets)+xlab("Time") + ylab("Unemployment Rate") + ggtitle("Residual for ETS")
```

2.  Residual of Mean forecast:

```{r}
mean_residual <- residuals(mean_forecast)
autoplot(mean_residual)+xlab("Time") + ylab("Unemployment Rate") + ggtitle("Residual for Mean")
```

3.  Residual of naive forecast:

```{r}
naive_residual <- residuals(naive_forecast)
autoplot(naive_residual)+xlab("Time") + ylab("Unemployment Rate") + ggtitle("Residual for Naive")
```

4.  Residual of STL Forecast:

```{r}
STL_residual <- residuals(data_stl)
autoplot(STL_residual)+xlab("Time") + ylab("Unemployment Rate") + ggtitle("Residual for STL")
```
